A general Nullstellensatz for generalized spaces

### 2019-05-05 paper-qcoh.tex
Beginning of the Horn subtopos case

The statement~\eqref{eq:nullstellensatz0} is not a geometric sequent. Therefore
it is not to be expected that it passes from~$\Set[\TT]$ to a
subtopos~$\Set[\TT']$ corresponding to a quotient theory~$\TT'$ of~$\TT$, and
indeed in general it does not. However, there is still a useful substitute,
which we formulate as Theorem~\ref{thm:nullstellensatz-horn}. This substitute
broadens the scope of the Nullstellensatz and is, whenever applicable, useful
for simplifying computations.

@@

\begin{thm}\label{thm:nullstellensatz-horn}
Let~$\TT$ be a geometric theory. Let~$\TT'$ be a quotient theory
of~$\TT$. Assume that the generic model~$U_\TT$ is a sheaf for the topology
on~$\Set[\TT]$ cutting out the subtopos~$\Set[\TT']$. Then the following
statement holds internally to~$\Set[\TT']$:
\[ \text{A geometric$^\star$ sequent~$\sigma$ with Horn consequent holds for~$U_{\TT'}$
  iff~$\ul{\TT}/U_\TT$ proves~$\sigma$.} \]
\end{thm}

@@

A situation in which the sheaf assumption of Theorem~\ref{thm:nullstellensatz-horn}
is satisfied is when~$\TT$ is a Horn theory and the topology cutting
out~$\Set[\TT']$ is subcanonical. For instance, this is the case if~$\TT$ is
the theory of rings and~$\Set[\TT']$ is one of several well-known toposes in
algebraic geometry such as the big Zariski topos, the big étale topos or the
big fppf topos.

\begin{rem}In the case that the~$\Set[\TT']$ is a dense subtopos
of~$\Set[\TT]$, Theorem~\ref{thm:nullstellensatz-horn} can be strengthened to
allow~$\bot$ as consequent, since in this case we have~$(\nabla\psi)
\Rightarrow \psi$ also for~$\psi \equiv \bot$.\end{rem}

@@

% XXX subtopos case

@@

Thanks to the version of the Nullstellensatz pertaining Horn consequents given
in Theorem~\ref{thm:nullstellensatz-horn}, these two arguments can also be
carried out in the classifying topos of local rings. It is instructive to
consider how we would have to argue if we wanted to use only the unmodified
Nullstellensatz:

>>>>

The statement~\eqref{eq:nullstellensatz0} is not a geometric sequent. Therefore
it is not to be expected that it passes from~$\Set[\TT]$ to a
subtopos~$\Set[\TT']$ corresponding to a quotient theory~$\TT'$ of~$\TT$, and
indeed in general it does not. However, there is still a useful substitute,
which we formulate as Corollary~\ref{corollary:nullstellensatz-horn}. This substitute
broadens the scope of the Nullstellensatz and is, whenever applicable, useful
for simplifying computations.

@@

\begin{cor}\label{corollary:nullstellensatz-horn}
Let~$\TT$ be a geometric theory. Let~$\TT'$ be a quotient theory
of~$\TT$. Assume that the generic model~$U_\TT$ is a sheaf for the topology
on~$\Set[\TT]$ cutting out the subtopos~$\Set[\TT']$. Then the following
statement holds internally to~$\Set[\TT']$:
\[ \text{A geometric$^\star$ sequent~$\sigma$ with Horn consequent holds for~$U_{\TT'}$
  iff~$\ul{\TT}/U_\TT$ proves~$\sigma$.} \]
\end{cor}

@@

A situation in which the sheaf assumption of Corollary~\ref{corollary:nullstellensatz-horn}
is satisfied is when~$\TT$ is a Horn theory and the topology cutting
out~$\Set[\TT']$ is subcanonical. For instance, this is the case if~$\TT$ is
the theory of rings and~$\Set[\TT']$ is one of several well-known toposes in
algebraic geometry such as the big Zariski topos, the big étale topos or the
big fppf topos.

\begin{rem}In the case that the~$\Set[\TT']$ is a dense subtopos
of~$\Set[\TT]$, Corollary~\ref{corollary:nullstellensatz-horn} can be strengthened to
allow~$\bot$ as consequent, since in this case we have~$(\nabla\psi)
\Rightarrow \psi$ also for~$\psi \equiv \bot$.\end{rem}

@@

% XXX The internal statement occurring in Theorem~\ref{thm:quasicoherent} is
% higher-order, therefore

\begin{cor}Let~$\TT'$ be a quotient theory of~$\TT$, not necessarily Horn.
Assume that the generic model~$U_\TT$ is a sheaf for the topology
on~$\Set[\TT]$ cutting out the subtopos~$\Set[\TT']$. Then, from the internal
point of view of~$\Set[\TT']$, the canonical map~$A \to (U_{\TT'})^{\Spec(A)}$
is an isomorphism of~$U_\TT$-algebras for every finitely
presented~$U_\TT$-algebra~$A$. (By~``$\Spec(A)$'', we mean the set
of~$U_\TT$-algebra homomorphisms from~$A$ to~$U_{\TT'}$.)
\end{cor}

@@

Thanks to the version of the Nullstellensatz pertaining Horn consequents given
in Corollary~\ref{corollary:nullstellensatz-horn}, these two arguments can also be
carried out in the classifying topos of local rings. It is instructive to
consider how we would have to argue if we wanted to use only the unmodified
Nullstellensatz:
